---
title: "Optimization Samples"
author: "Yale Waller"
date: "October 21, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Setup
```{r, echo=FALSE}
#install.packages("optrees")
#install.packages("igraph")
#install.packages("qgraph")
require(optrees)
require(igraph)
require(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

#Section 1
``` {r}
n<-1000
d<-runif(n*n)
d[d<0.8]<-NA
d<-matrix(d,nrow=n,ncol=n)
diag(d)<-NA
d[upper.tri(d)]=t(d)[upper.tri(d)]
str(d)
nrow(d)

sparsify<-function (x) {
#progressbar if we want it
#fog<-txtProgressBar(min = 0, max = nrow(x), style = 3)

#actual function begins here
nodes<-seq(1,nrow(x),1)
diag(x)<-NA
    qqq <- NULL
        for (i in seq(along = nodes)) {
		#setTxtProgressBar(fog, i)
            for (j in i:length(nodes)) {
                if (!is.na(x[i, j]) && x[i, j] != Inf) {
                  qqq <- rbind(qqq, c(i, j, x[i, j]))
                }
            }
        }
      colnames(qqq) <- c("head", "tail", "weight")
	return(qqq)
	close(pb)
    }

d2<-sparsify(d)
str(d2)
head(d2)
```

#Sections 2 & 3
``` {r}
n<-250
x<-round(runif(n)*1000)
y<-round(runif(n)*1000)
adjmat<-matrix(,n,n)
for (i in 1:nrow(adjmat)) {
	for (q in 1:ncol(adjmat)) {
	adjmat[i,q]<- sqrt((y[q]-y[i])^2 + (x[q]-x[i])^2)
	}
}
ds<-sparsify(adjmat)
#str(ds)
#head(ds)

ds.mst<-msTreePrim(1:n,ds)
str(ds.mst)
arclist<-NULL
```

```{r}
plot.mst<-function (arclist) {
	for (i in 1:length(arclist[,1])) {
	segments(x0=x[arclist[i,1]], y0=y[arclist[i,1]], x1=x[arclist[i,2]],y1=y[arclist[i,2]])
	}

}
plot(x,y,pch=16)
plot.mst(ds.mst$tree.arcs)
```

#Section 4
1. The probability of a message falling into enemy hands is the cost. The agents are the nodes, this is a minimum spanning tree since they want to minimize the probability.
2. The same way we have previously, by creating a sparse matrix with three columns, agents as destination and source nodes, probabilities as costs.
3. The answer to 'which algorithm' depends greatly upon the setup of the 'agents', Prim's would be better if the graph is dense, with more edges than vertices. That might be the case if all agents are interconnected. However if the graph is sparse, say these are agent hubs in Russia, it might be better to use Kruskal's. However, I would select Prim's given the likelihood of the former.
4. The computational efficiency is O(|V|^2), assuming an adjacency matrix as the input.

#Section 5.1-5.8
``` {r}
#Edge weights
Duration<-c(90,15,5,20,21,25,14,28,30,45)

#first pass of Bellman-Ford
arcs<-matrix(c(1,2,-90, 2,3,-15, 7,4,-14, 4,5,-20, 1,6,-90,
3,7,-5, 6,7,-25, 4,8,-20, 1,9,-90, 4,10,-20, 9,10,-30), ncol=3, byrow=TRUE)
nodes<-1:10
minspan<-getShortestPathTree(nodes, arcs, algorithm="Bellman-Ford", directed=TRUE)

#second transposed pass of Bellman-Ford
arcs<-matrix(c(2,1,-15, 3,2,-5, 4,7,-20, 5,4,-21, 6,1,-25,
7,3,-14, 7,6,-14, 8,4,-28, 9,1,-30, 10,4,-45, 10,9,-45, 11,5,0, 11,8,0, 11,10,0), ncol=3, byrow=TRUE)
nodes<-1:11
minspan2<-getShortestPathTree(nodes, arcs, algorithm="Bellman-Ford", directed=TRUE, source.node=11)

#Construct the table with all the necessary columns
NodeLabels<-c('A','B','C','D','E','F','G','H','I','J')

#EarliestStart
EarliestStart<-abs(minspan$distances)

#EarliestFinish
EarliestFinish<-abs(minspan$distances)+Duration

#LatestFinish
addmax<-max(EarliestFinish)
LatestFinish = rep(0, length(Duration))
for (i in 1:length(Duration)) {
	LatestFinish[i] = minspan2$distances[i] + addmax
}

#LatestStart
LatestStart<-rep(0, length(Duration))
for (i in 1:length(Duration)){
	LatestStart[i] = LatestFinish[i]-Duration[i]
}

#Slack
Slack<-LatestFinish-EarliestFinish

table<-data.frame(NodeLabels,Duration,EarliestStart,EarliestFinish,LatestStart,LatestFinish,Slack)
table

print(paste("The Earliest project completion date is ", as.Date('2017-11-01') + max(EarliestFinish)))

```

#5.9
Below are the Nodes that are first part of the critical path, and those with some flexibility, respectively.
```{r}
NodeLabels[which(Slack==0)] #part of the critical path
NodeLabels[which(Slack!=0)] #flexibility
```

#5.10, Charting
```{r}
barplot(rbind(LatestFinish-Duration,Duration),horiz=TRUE,col=c("white","gray"),
border=0,names.arg=NodeLabels,las=1,main="Estimated Project Completion Times")

```


